---
title: "Results"
author: "Emma Sims, Jonathan Trattner, and S. Mason Garrison"
date: "`r format(Sys.time(), '%B %d %Y')`"
output:
  html_document: default
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup,include=FALSE}
## TODO: add distinct(extended_id, .keep_all = TRUE) # default is to keep the first row


knitr::opts_chunk$set(echo = TRUE,error=TRUE)

# set seed
set.seed(20200804)

## Installs library if missing

if (!require("tidyverse")) install.packages("tidyverse")
if (!require("devtools")) install.packages("devtools")
if (!require("remotes")) install.packages("remotes")
#if (!require("NlsyLinks")) install.packages("NlsyLinks")
#if (!require("discord")) install.packages("discord")
if (!require("janitor")) install.packages("janitor")
if (!require("lm.beta")) install.packages("lm.beta")
if (!require("matrixStats")) install.packages("matrixStats")
if (!require("tictoc")) install.packages("tictoc")
if (!require("glue")) install.packages("glue")
if (!require("psych")) install.packages("psych")
if (!require("foreign")) install.packages("foreign")
if(!require("ggplot2")) install.packages("ggplot2")
if(!require("parameters")) install.packages("parameters")
if(!require("gt")) install.packages("gt")

## Installs most recent dev version
remotes::install_github(repo="nlsy-links/NlsyLinks")
devtools::install_github('R-Computing-Lab/discord') 

if (!require("dplyr")) install.packages("dplyr")

library(tidyverse)
library(parameters)
library(devtools) 
library(remotes)
library(NlsyLinks)
library(gtsummary)
library(discord)
library(janitor)
library(lm.beta)
library(matrixStats)
library(tictoc)
library(glue)
library(psych)
library(foreign)
library(dplyr)
library(ggplot2)
library(gt)



# functions

## good workaround for stupid default setting; runningg mean function but we are overriding default
new_mean=function(x){
  mean(x,na.rm=TRUE)
}

prettify_regression_results <- function(regression_object,
                                        intercept=TRUE,
                                        standardize=FALSE,
					digits=3) {
										
temp <-  regression_object %>%
            gtsummary::tbl_regression(intercept=intercept, 
                                      pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = digits),
                                      estimate_fun = ~ gtsummary::style_sigfig(.x, digits = digits)
            ) %>%
        gtsummary::modify_header(
            statistic ~ "**t-statistic**", p.value ~ "**p-value**"
        ) %>%
        gtsummary::add_glance_source_note(
            label = list(statistic ~ "F-statistic",
                         df  ~ "DF1",
                         df.residual  ~ "DF2"),
            include = c(r.squared, statistic, df, df.residual, p.value, nobs)
        )

if(standardize){
temp_stnd <-      regression_object %>%
        gtsummary::tbl_regression(intercept=intercept, 
                                  estimate_fun = ~ gtsummary::style_sigfig(.x, digits = digits),
                                  pvalue_fun = ~ gtsummary::style_pvalue(.x, digits = digits),
                                  tidy_fun = gtsummary::tidy_standardize,
								  conf.int=FALSE) %>%
        gtsummary::modify_header(
            estimate ~ "**β**"
        ) 

temp <-
  tbl_merge(
    list(temp_stnd, temp),
    tab_spanner = c("**Standardized**", "**Unstandardized**")
  )
}
temp  
}
```

# Data

## Import

### Depression

```{r df_dep, echo=FALSE}

source('scripts/depression.R')

varlabels <- c("CES-D - POOR APPETITE XRND",
  "CES-D - COULD NOT SHAKE BLUES XRND",
  "CES-D - TROUBL KEEPING MIND ON TASKS XRND",
  "CES-D - DEPRESSED XRND",
  "CES-D - EVERYTHING TOOK EXTRA EFFORT XRND",
  "CES-D - RESTLESS SLEEP XRND",
  "CES-D - FELT LONELY XRND",
  "CES-D - SAD XRND",
  "CES-D - COULD NOT GET GOING XRND",
  "H40 7-ITEM CES-D SCORE XRND",
  "SF-12 - EMOTIONAL PRBS LIMIT ACCMPLSHMT? XRND",
  "SF-12 - EMOTIONAL PRBS MADE LESS CARFUL? XRND",
  "SF-12 - HLTH PRBS HINDER SOC ACTIV? XRND",
  "CCR-DR DIAGNS EMTNL,NRVS,PSYC PRBS? XRND",
  "CCR-DATE DR DIAG EMTN,NRVS,PSYC PRB XRND",
  "CCR-DATE DR DIAG EMTN,NRVS,PSYC PRB XRND",
  "CCR-EMTN,NRVS,PSYC PROBS LST 12 MO? XRND",
  "CCR-DEPRESN,EXCESS WORRY,NERVS PRB XRND",
  "H50 7-ITEM CES-D SCORE XRND",
  "CESD -  POOR APPETITE XRND",
  "CESD - COULD NOT SHAKE BLUES XRND",
  "CESD - TROUBLE KEEPING MIND ON TASK XRND",
  "CESD - DEPRESSION XRND",
  "CESD - EVERYTHING TAKING EXTRA EFFORT XRND",
  "CESD - RESTLESS SLEEP XRND",
  "CESD - FEELING LONELY XRND",
  "CESD -  FEELING SAD XRND",
  "CESD - COULD NOT GET GOING XRND",
  "SF12 - EMOTIONAL PRBS LIMIT ACCMPLSHMT? XRND",
  "SF12 - EMOTIONAL PRBS MADE LESS CARFUL? XRND",
  "SF12 - CALM/PEACEFUL PAST 4 WKS? XRND",
  "SF12 - HAVE ENERGY PAST 4 WKS? XRND",
  "SF12 - DOWNHRTD OR BLUE PAST 4 WKS? XRND",
  "SF12 - HLTH PRBS HINDER SOC ACTIV? XRND",
  "DOCTOR EVER DIAGNOSED R SUFFERING FROM DEPRESSION XRND",
  "DATE RS DEPRESSION DIAGNOSED XRND",
  "DATE RS DEPRESSION DIAGNOSED XRND",
  "R SUFFERED FROM DEPRESSION IN LAST 12 MOS XRND",
  "PSYCHIATRIC DIAGNOSIS IN 40+ HEALTH MOD XRND",
  "DR EVER SAID R HAD EMO/NERV/PSYC PBLS OT THAN DEP XRND",
  "DATE R EMO/NERV/PSYC PBLS DIAGNOSED XRND",
  "DATE R EMO/NERV/PSYC PBLS DIAGNOSED XRND",
  "R HAD ANY EMO/NERV/PSYC PROBLEMS IN LAST 12 MOS XRND",
  "ID# (1-12686) 79",
  "HLTH CONSOLIDATION CODE 5-D 79",
  "SAMPLE ID  79 INT",
  "CES-D-BOTHERED BY THINGS 92",
  "CES-D-POOR APPETITE 92",
  "CES-D-UNABLE TO SHAKE BLUES 92",
  "CES-D-FELT GOOD AS OTHR PEOPLE 92",
  "CES-D-TRBLE KEEPING MIND ON TASKS 92",
  "CES-D-DEPRESSED 92",
  "CES-D-THINGS TOOK EXTRA EFFORT 92",
  "CES-D-HOPEFUL 92",
  "CES-D-A FAILURE 92",
  "CES-D-FEARFUL 92",
  "CES-D-RESTLESS SLEEP 92",
  "CES-D-HAPPY 92",
  "CES-D-LESS TALKATIVE THAN USUAL 92",
  "CES-D-LONELY 92",
  "CES-D-OTHERS WERE UNFRIENDLY 92",
  "CES-D-ENJOYED LIFE 92",
  "CES-D-CRYING SPELLS 92",
  "CES-D-SAD 92",
  "CES-D-DISLIKED BY OTHERS 92",
  "CES-D-COULD NOT GET GOING 92",
  "20-ITEM CES-D SCORE 92",
  "FLAG FOR 20-ITEM CES-D SCORE 92",
  "7-ITEM CES-D SCORE 92",
  "CES-D - POOR APPETITE 94",
  "CES-D-TRBLE KPNG MIND ON TASKS 94",
  "CES-D - DEPRESSED 94",
  "CES-D - TOOK EXTRA EFFORT 94",
  "CES-D - RESTLESS SLEEP 94",
  "CES-D - SAD 94",
  "CES-D - COULD NOT GET GOING 94",
  "7-ITEM CES-D SCORE 94",
  "R SATISFACTION WITH LIFE 2014",
  "R SATISFACTION WITH LIFE 2016"
)


# Use qnames rather than rnums

qnames = function(data) {
  names(data) <- c("H40_CESD_01",
    "H40_CESD_02",
    "H40_CESD_03",
    "H40_CESD_04",
    "H40_CESD_05",
    "H40_CESD_06",
    "H40_CESD_07",
    "H40_CESD_08",
    "H40_CESD_09",
    "H40_CESD_SCORE_7_ITEM",
    "H40_SF12_5",
    "H40_SF12_5B",
    "H40_SF12_8",
    "H40_CHRC_8",
    "H40_CHRC_8A_M",
    "H40_CHRC_8A_Y",
    "H40_CHRC_8B",
    "H40_CHRC_10K",
    "H50_CESD_SCORE_7_ITEM",
    "H50_CESD_01",
    "H50_CESD_02",
    "H50_CESD_03",
    "H50_CESD_04",
    "H50_CESD_05",
    "H50_CESD_06",
    "H50_CESD_07",
    "H50_CESD_08",
    "H50_CESD_09",
    "H50_SF12_5",
    "H50_SF12_5B",
    "H50_SF12_7",
    "H50_SF12_7B",
    "H50_SF12_7C",
    "H50_SF12_8",
    "H50_CHRC_7B",
    "H50_CHRC_7C_M",
    "H50_CHRC_7C_Y",
    "H50_CHRC_7D",
    "H50_CHRC_CHK6",
    "H50_CHRC_8",
    "H50_CHRC_8A_M",
    "H50_CHRC_8A_Y",
    "H50_CHRC_8B",
    "CASEID",
    "HEALTH_CODE_1979",
    "SAMPLE_ID",
    "CESD_00_1992",
    "CESD_01_1992",
    "CESD_02_1992",
    "CESD_02A_1992",
    "CESD_03_1992",
    "CESD_04_1992",
    "CESD_05_1992",
    "CESD_05A_1992",
    "CESD_05B_1992",
    "CESD_05C_1992",
    "CESD_06_1992",
    "CESD_06A_1992",
    "CESD_06B_1992",
    "CESD_07_1992",
    "CESD_07A_1992",
    "CESD_07B_1992",
    "CESD_07C_1992",
    "CESD_08_1992",
    "CESD_08A_1992",
    "CESD_09_1992",
    "CESD_SCORE_20_ITEM_1992",
    "CESD_FLAG_1992",
    "CESD_SCORE_7_ITEM_1992",
    "CESD_01_1994",
    "CESD_03_1994",
    "CESD_04_1994",
    "CESD_05_1994",
    "CESD_06_1994",
    "CESD_08_1994",
    "CESD_09_1994",
    "CESD_SCORE_7_ITEM_1994",
    "LIFE-SATISFACTION_2014",
    "LIFE-SATISFACTION_2016")
  return(data)
}



df_dep_gen1 <- qnames(vallabels(new_data))

remove(new_data)


```

### Crime

```{r data_crime, echo=FALSE}
source('scripts/crime.R')

varlabels <- c("ID# (1-12686) 79",
  "SAMPLE ID  79 INT",
  "RACL/ETHNIC COHORT /SCRNR 79",
  "SEX OF R 79",
  "ILL ACT RUN AWAY PAST YR (AGE <18) 80",
  "ILL ACT INTENTION DAMAGED PROP PAST 80",
  "ILL ACT FOUGHT @ SCHOOL/WRK PAST YR 80",
  "ILL ACT SHOPLIFTED PAST YR 80",
  "ILL ACT STEEL PAST YR (<$50) 80",
  "ILL ACT STEEL PAST YR (>$50) 80",
  "ILL ACT USED FORCE OBTAIN THINGS YR 80",
  "ILL ACT # SERIOUS THREAT HIT PAST 80",
  "ILL ACT ATTACK W/INTNT INJR/KILL PAST 80",
  "ILL ACT SOLD MARJ PAST YR 80",
  "ILL ACT SOLD HARD DRUGS PAST YR 80",
  "ILL ACT ATTEMPTED TO 80",
  "ILL ACT TAKE AUTO W/O PERM PAST YR 80",
  "ILL ACT BROKEN INTO A BUILDING PAST 80",
  "ILL ACT KNOW SOLD/HELD STOLEN PAST 80",
  "ILL ACT AID GAMBLING OPER PAST YR 80",
  "STOP BY POLICE O/THAN MIN TRAFIC OFF 80",
  "TIMES STOP BY POLICE MINOR 80",
  "TIMES STOP POLICE EXCLD MINOR PAST YR 80",
  "AGE OF R 1ST TIME STOPPED BY POLICE 80",
  "EVER CHARGED W/ILGL ACT 80",
  "TIMES CHARGE W/ILGL ACT EXCLD MINOR 80",
  "TIMES CHARGE W/ILGL ACT EXCLD MINOR 80",
  "MONTH M-RCNT ILGL ACT CHARGE 80",
  "YR M-RCNT ILGL ACT CHARGE 80",
  "AGE @ TIME 1ST ILGL ACT CHARGE 80",
  "EVER CHARGED ILGL ACT ADLT COURT 80",
  "EVER CONVICTED ON ILGL ACT CHARGES 80",
  "TIMES CONVICTED ILGL ACT EXCLD MINOR 80",
  "AGE @ TIME 1ST ILGL ACT CONVICTION 80",
  "MONTH M-RCNT ILGL ACT CONVICTION 80",
  "YR M-RCNT ILGL ACT CONVICTION 80",
  "ILL ACT ASSAULT 80",
  "ILL ACT ROBBERY 80",
  "ILL ACT THEFT 80",
  "ILL ACT THEFT BY DECEPTION 80",
  "ILL ACT STOLEN PROP 80",
  "ILL ACT DESTRUCTION OF PROP 80",
  "ILL ACT OTHER PROP OFFENSE 80",
  "ILL ACT GAMBLING 80",
  "ILL ACT COMMERCIAL VICE 80",
  "ILL ACT POSSESSION OF MARJ/HASH 80",
  "ILL ACT SELL MARJ/HASH 80",
  "ILL ACT POSSESSION OTHER DRUGS 80",
  "ILL ACT SALE/MANUF ILLICIT DRUGS 80",
  "ILL ACT MAJOR TRAFIC OFFENSE 80",
  "ILL ACT STATUS OFFENSE 80",
  "ILL ACT OTHER 80",
  "EVER CONVICTED ILGL ACT ADLT COURT 80",
  "EVER REFUSED COURT-RELATED COUNSELING 80",
  "TIMES REFER COURT-RELATED COUNSELING 80",
  "AGE @ TIME 1ST COURT-REL COUNSELING 80",
  "MONTH M-RCNT COURT-REL COUNSELING END 80",
  "YR M-RCNT COURT-REL COUNSELING END 80",
  "EVER BEEN ON PROBATION 80",
  "TIMES ON PROBATION 80",
  "DATE M-RCNT PROBATION PRD END 80",
  "YR M-RCNT PROBATION PRD END 80",
  "EVER SNTNCD ANY CORP INSTITUTN 80",
  "TIMES SENT TO YTH CORP INSTITUTN 80",
  "TIMES SENT TO ADLT CORP INSTITUTN 80",
  "MONTH M-RCNT RLSE CORP INSTITUTN 80",
  "YR M-RCNT RLSE CORP INSTITUTN 80",
  "REASON FOR NONINT 80",
  "REASON FOR NONINT 81",
  "REASON FOR NONINT 82",
  "REASON FOR NONINT 83",
  "REASON FOR NONINT 84",
  "REASON FOR NONINT 85",
  "REASON FOR NONINT 86",
  "REASON FOR NONINT 87",
  "REASON FOR NONINT 88",
  "REASON FOR NONINT 89",
  "REASON FOR NONINT 90",
  "REASON FOR NONINT 91",
  "REASON FOR NONINT 92",
  "REASONS FOR NON-INT 93",
  "ARRESTED, IN POLICE TROUBLE 94",
  "REASONS FOR NON-INT 94",
  "REASONS FOR NON-INT 96",
  "REASONS FOR NON-INT 1998",
  "REASONS FOR NON-INT 2000",
  "REASONS FOR NON-INT 2002",
  "REASONS FOR NON-INT 2004",
  "REASONS FOR NON-INT 2006",
  "REASONS FOR NON-INT 2008",
  "REASONS FOR NON-INT 2010",
  "REASONS FOR NON-INT 2012",
  "REASONS FOR NON-INT 2014",
  "REASONS FOR NON-INT 2016"
)

# Use qnames rather than rnums

qnames = function(data) {
  names(data) <- c("CASEID",
     "SAMPLE_ID",	#	  "SAMPLE ID  79 INT",
    "RACE",	#	  "RACL/ETHNIC COHORT /SCRNR 79",
    "SEX",	#	  "SEX OF R 79",
    "DELIN_1_1980",	  #	  "ILL ACT RUN AWAY PAST YR (AGE <18) 80",
    "DELIN_4_1980",	  #	  "ILL ACT INTENTION DAMAGED PROP PAST 80",
    "DELIN_5_1980",	  #	  "ILL ACT FOUGHT @ SCHOOL/WRK PAST YR 80",
    "DELIN_6_1980",	  #	  "ILL ACT SHOPLIFTED PAST YR 80",
    "DELIN_7_1980",	  #	  "ILL ACT STEEL PAST YR (<$50) 80",
    "DELIN_8_1980",	  #	  "ILL ACT STEEL PAST YR (>$50) 80",
    "DELIN_9_1980",	  #	  "ILL ACT USED FORCE OBTAIN THINGS YR 80",
    "DELIN_10_1980",	#	  "ILL ACT # SERIOUS THREAT HIT PAST 80",
    "DELIN_11_1980",	#	  "ILL ACT ATTACK W/INTNT INJR/KILL PAST 80",
    "DELIN_14_1980",	#	  "ILL ACT SOLD MARJ PAST YR 80",
    "DELIN_15_1980",	#	  "ILL ACT SOLD HARD DRUGS PAST YR 80",
    "DELIN_16_1980",	#	  "ILL ACT ATTEMPTED TO 80",
    "DELIN_17_1980",	#	  "ILL ACT TAKE AUTO W/O PERM PAST YR 80",
    "DELIN_18_1980",	#	  "ILL ACT BROKEN INTO A BUILDING PAST 80",
    "DELIN_19_1980",	#	  "ILL ACT KNOW SOLD/HELD STOLEN PAST 80",
    "DELIN_20_1980",	#	  "ILL ACT AID GAMBLING OPER PAST YR 80",
    "POLICE_1_1980",	#	  "STOP BY POLICE O/THAN MIN TRAFIC OFF 80",
    "POLICE_1A_1980",	#	  "TIMES STOP BY POLICE MINOR 80",
    "POLICE_1B_1980",	#	  "TIMES STOP POLICE EXCLD MINOR PAST YR 80",
    "POLICE_1C_1980",	#	  "AGE OF R 1ST TIME STOPPED BY POLICE 80",
    "POLICE_2_1980",	#	  "EVER CHARGED W/ILGL ACT 80",
    "POLICE_2A_1980",	#	  "TIMES CHARGE W/ILGL ACT EXCLD MINOR 80",
    "POLICE_2B_1980",	#	  "TIMES CHARGE W/ILGL ACT EXCLD MINOR 80",
    "POLICE_2C_M_1980",	#	  "MONTH M-RCNT ILGL ACT CHARGE 80",
    "POLICE_2C_Y_1980",	#	  "YR M-RCNT ILGL ACT CHARGE 80",
    "POLICE_2D_1980",	#	  "AGE @ TIME 1ST ILGL ACT CHARGE 80",
    "POLICE_2E_1980",	#	  "EVER CHARGED ILGL ACT ADLT COURT 80",
    "POLICE_3_1980",	#	  "EVER CONVICTED ON ILGL ACT CHARGES 80",
    "POLICE_3A_1980",	#	  "TIMES CONVICTED ILGL ACT EXCLD MINOR 80",
    "POLICE_3B_1980",	#	  "AGE @ TIME 1ST ILGL ACT CONVICTION 80",
    "POLICE_3C_M_1980",	#	  "MONTH M-RCNT ILGL ACT CONVICTION 80",
    "POLICE_3C_Y_1980",	#	  "YR M-RCNT ILGL ACT CONVICTION 80",
    "POLICE_3D_01_1980",	#	  "ILL ACT ASSAULT 80",
    "POLICE_3D_02_1980",	#	  "ILL ACT ROBBERY 80",
    "POLICE_3D_03_1980",	#	  "ILL ACT THEFT 80",
    "POLICE_3D_04_1980",	#	  "ILL ACT THEFT BY DECEPTION 80",
    "POLICE_3D_05_1980",	#	  "ILL ACT STOLEN PROP 80",
    "POLICE_3D_06_1980",	#	  "ILL ACT DESTRUCTION OF PROP 80",
    "POLICE_3D_07_1980",	#	  "ILL ACT OTHER PROP OFFENSE 80",
    "POLICE_3D_08_1980",	#	  "ILL ACT GAMBLING 80",
    "POLICE_3D_09_1980",	#	  "ILL ACT COMMERCIAL VICE 80",
    "POLICE_3D_10_1980",	#	  "ILL ACT POSSESSION OF MARJ/HASH 80",
    "POLICE_3D_11_1980",	#	  "ILL ACT SELL MARJ/HASH 80",
    "POLICE_3D_12_1980",	#	  "ILL ACT POSSESSION OTHER DRUGS 80",
    "POLICE_3D_13_1980",	#	  "ILL ACT SALE/MANUF ILLICIT DRUGS 80",
    "POLICE_3D_14_1980",	#	  "ILL ACT MAJOR TRAFIC OFFENSE 80",
    "POLICE_3D_16_1980",	#	  "ILL ACT STATUS OFFENSE 80",
    "POLICE_3D_17_1980",	#	  "ILL ACT OTHER 80",
    "POLICE_3E_1980",   #	  "EVER CONVICTED ILGL ACT ADLT COURT 80",
    "POLICE_4_1980",    #	  "EVER REFUSED COURT-RELATED COUNSELING 80",
    "POLICE_4A_1980",   #	  "TIMES REFER COURT-RELATED COUNSELING 80",
    "POLICE_4B_1980",   #	  "AGE @ TIME 1ST COURT-REL COUNSELING 80",
    "POLICE_4C_M_1980",	#	  "MONTH M-RCNT COURT-REL COUNSELING END 80",
    "POLICE_4C_Y_1980",	#	  "YR M-RCNT COURT-REL COUNSELING END 80",
    "POLICE_6_1980",    #	  "EVER BEEN ON PROBATION 80",
    "POLICE_6A_1980",	  #	  "TIMES ON PROBATION 80",
    "POLICE_6B_M_1980",	#	  "DATE M-RCNT PROBATION PRD END 80",
    "POLICE_6B_Y_1980",	#	  "YR M-RCNT PROBATION PRD END 80",
    "POLICE_7_1980",    #	  "EVER SNTNCD ANY CORP INSTITUTN 80",
    "POLICE_7A_1980",   #	  "TIMES SENT TO YTH CORP INSTITUTN 80",
    "POLICE_7B_1980",   #	  "TIMES SENT TO ADLT CORP INSTITUTN 80",
    "POLICE_7C_M_1980",	#	  "MONTH M-RCNT RLSE CORP INSTITUTN 80",
    "POLICE_7C_Y_1980",	#	  "YR M-RCNT RLSE CORP INSTITUTN 80",
    "RNI_1980",	        #	  "REASON FOR NONINT 80",
    "RNI_1981",	        #	  "REASON FOR NONINT 81",
    "RNI_1982",         #	  "REASON FOR NONINT 82",
    "RNI_1983",         #	  "REASON FOR NONINT 83",
    "RNI_1984",         #	  "REASON FOR NONINT 84",
    "RNI_1985",         #	  "REASON FOR NONINT 85",
    "RNI_1986",         #	  "REASON FOR NONINT 86",
    "RNI_1987",         #	  "REASON FOR NONINT 87",
    "RNI_1988",         #	  "REASON FOR NONINT 88",     
    "RNI_1989",         #	  "REASON FOR NONINT 89",
    "RNI_1990",         #	  "REASON FOR NONINT 90",
    "RNI_1991",         #	  "REASON FOR NONINT 91",
    "RNI_1992",         #	  "REASON FOR NONINT 92",
    "RNI_1993",         #	  "REASONS FOR NON-INT 93",
    "ARREST_7N_1994",   #	  "ARRESTED, IN POLICE TROUBLE 94",
    "RNI_1994",         #	  "REASONS FOR NON-INT 94",
    "RNI_1996",         #	  "REASONS FOR NON-INT 96",
    "RNI_1998",         #	  "REASONS FOR NON-INT 1998",
    "RNI_2000",         #	  "REASONS FOR NON-INT 2000",
    "RNI_2002",         #	  "REASONS FOR NON-INT 2002",
    "RNI_2004",         #	  "REASONS FOR NON-INT 2004",
    "RNI_2006",         #	  "REASONS FOR NON-INT 2006",
    "RNI_2008",         #	  "REASONS FOR NON-INT 2008",
    "RNI_2010",         #	  "REASONS FOR NON-INT 2010",
    "RNI_2012",         #	  "REASONS FOR NON-INT 2012",
    "RNI_2014",         #	  "REASONS FOR NON-INT 2014",
    "RNI_2016")         #	  "REASONS FOR NON-INT 2016"

  return(data)
}


df_crime_gen1 <- qnames(vallabels(new_data))


remove(new_data)

```

### age

```{r people_age, echo=FALSE}

#How old is everyone? 
source("scripts/age.R")

varlabels <- c("ID# (1-12686) 79",
  "DATE OF BIRTH - MONTH 79",
  "DATE OF BIRTH - YR 79",
  "AGE OF R 79",
  "SAMPLE ID  79 INT",
  "RACL/ETHNIC COHORT /SCRNR 79",
  "SEX OF R 79"
)


# Use qnames rather than rnums

qnames = function(data) {
  names(data) <- c("CASEID",
    "M_1979",
    "Y_1979",
    "AGE_1979",
    "SAMPLE_ID",
    "RACE_78SCRN",
    "SEX_1979")
  return(data)
}


#********************************************************************************************************


# Remove the '#' before the following lines to rename variables using Qnames instead of Reference Numbers
df_age <- qnames(vallabels(df_age))



```

## Clean

```{r data_cleaning, echo=FALSE}

# is average delinquency score; so that missing data doesn't deflate score
df_crime_gen1 <- 
  df_crime_gen1 %>% 
  mutate(
    SubjectTag=CASEID*100,
    dq_sum = rowSums(select(.,starts_with("DELIN_")),na.rm=TRUE),
    DELIN_AVERAGE = rowMeans(select(.,starts_with("DELIN_")),na.rm=TRUE)#,
#DELIN_Median = rowMedians(select(., starts_with("DELIN_")),na.rm=TRUE)
)


# creating smaller dataframe to only include variables of interest
# recoding Race and Sex ; todo, name race to minority

df_data_gen1 <- df_crime_gen1 %>%
  select(
    SubjectTag,
    CASEID,
    RACE,
    SEX,
    DELIN_AVERAGE
  ) %>%
  mutate(
      # effectively race is 0 for white; and 1 for non-white
    MINORITY = case_when(
      RACE == 3 ~ "nonblackhispanic",
      RACE %in% c(1, 2) ~ "blackorhispanic"       #black or hisp
    ),
    # have female be 0 and male be 1
    SEX = case_when(
      SEX == 2 ~ 0,
      TRUE ~ 1
    )
  ) %>% 
  bind_cols(
    CESD7_1994 = df_dep_gen1$CESD_SCORE_7_ITEM_1994,
    CESD20_1992 = as.numeric(df_dep_gen1$CESD_SCORE_20_ITEM_1992),
    CESD7_1992 = df_dep_gen1$CESD_SCORE_7_ITEM_1992,
    AGE_1979 = df_age$AGE_1979
  )


df_data_gen1 %>%
  select(-c(SEX,
            MINORITY,
            RACE,
            SubjectTag,
            CASEID)) %>%
  cor(use = "pairwise.complete")

# convert sex into factor
df_data_gen1 <- df_data_gen1 %>%  mutate(
    SEX = as.factor(SEX),
    RACE = as.factor(RACE),
    MINORITY = as.factor(MINORITY)
  )



#summary(df_data_gen1)


stardadized_by_gender=FALSE
# standardize by gender
if(stardadized_by_gender){
df_data_gen1 <- group_by(df_data_gen1, SEX) %>% 
  mutate(DELIN_AVERAGE = as.numeric(scale(DELIN_AVERAGE)),
         CESD7_1992 = as.numeric(scale(CESD7_1992)),
         CESD20_1992 = as.numeric(scale(CESD20_1992)),
         CESD7_1994 = as.numeric(scale(CESD7_1994))) %>%
ungroup()
}

# Group the data by sex and calculate mean, standard deviation, and sample size for different variables
df_data_SEX <- df_data_gen1 %>% 
  group_by(SEX) %>% 
  summarize(mean_Delin_Avg = mean(DELIN_AVERAGE, na.rm = TRUE), # calculates the mean for the DELIN_AVERAGE variable
            sd_Delin_Avg = sd(DELIN_AVERAGE, na.rm = TRUE),   # calculates the standard deviation for the DELIN_AVERAGE variable
            n_Delin_Avg = n(), # calculates the sample size for the DELIN_AVERAGE variable
            mean_CESD7_1992 = mean(CESD7_1992, na.rm = TRUE), # calculates the mean for the CESD7_1992 variable
            sd_CESD7_1992 = sd(CESD7_1992, na.rm = TRUE),  # calculates the standard deviation for the CESD7_1992 variable
            n_CESD7_1992 = n(), # calculates the sample size for the CESD7_1992 variable
            mean_CESD7_1994 = mean(CESD7_1994, na.rm = TRUE), # calculates the mean for the CESD7_1994 variable
            sd_CESD7_1994 = sd(CESD7_1994, na.rm = TRUE),  # calculates the standard deviation for the CESD7_1994 variable
            n_CESD7_1994 = n(), # calculates the sample size for the CESD7_1994 variable
            mean_CESD20_1992 = mean(CESD20_1992, na.rm = TRUE), # calculates the mean for the CESD20_1992 variable
            sd_CESD20_1992 = sd(CESD20_1992, na.rm = TRUE),  # calculates the standard deviation for the CESD20_1992 variable
            n_CESD20_1992 = n()) %>%   # calculates the sample size for the CESD20_1992 variable
  ungroup() # remove grouping from the data frame

# Display the resulting data frame as a nice table using the gt package
gt(df_data_SEX, title = "Depression and Delinquency means divided by Gender")

#age

# Group the data by age and calculate mean, standard deviation, and sample size for different variables
df_data_AGE <- df_data_gen1 %>% 
  group_by(AGE_1979) %>% 
  summarize(mean_Delin_Avg = mean(DELIN_AVERAGE, na.rm = TRUE), # calculates the mean for the DELIN_AVERAGE variable
            sd_Delin_Avg = sd(DELIN_AVERAGE, na.rm = TRUE),   # calculates the standard deviation for the DELIN_AVERAGE variable
            n_Delin_Avg = n(), # calculates the sample size for the DELIN_AVERAGE variable
            mean_CESD7_1992 = mean(CESD7_1992, na.rm = TRUE),
            sd_CESD7_1992 = sd(CESD7_1992, na.rm = TRUE),  
            n_CESD7_1992 = n(),
            mean_CESD7_1994 = mean(CESD7_1994, na.rm = TRUE),
            sd_CESD7_1994 = sd(CESD7_1994, na.rm = TRUE),  
            n_CESD7_1994 = n(),
            mean_CESD20_1992 = mean(CESD20_1992, na.rm = TRUE),
            sd_CESD20_1992 = sd(CESD20_1992, na.rm = TRUE),  
            n_CESD20_1992 = n()) %>%   
  ungroup()

gt(df_data_AGE, title = "Depression and Delinquency means divided by Age")

#race
library(tidyverse)

# Group the data by minority status and calculate mean, standard deviation, and sample size for different variables
df_data_RACE <- df_data_gen1 %>% 
  group_by(MINORITY) %>% 
  summarize(mean_Delin_Avg = mean(DELIN_AVERAGE, na.rm = TRUE), # calculates the mean for the DELIN_AVERAGE variable
            sd_Delin_Avg = sd(DELIN_AVERAGE, na.rm = TRUE),   # calculates the standard deviation for the DELIN_AVERAGE variable
            n_Delin_Avg = n(), # calculates the sample size for the DELIN_AVERAGE variable
            mean_CESD7_1992 = mean(CESD7_1992, na.rm = TRUE), # calculates the mean for the CESD7_1992 variable
            sd_CESD7_1992 = sd(CESD7_1992, na.rm = TRUE),  # calculates the standard deviation for the CESD7_1992 variable
            n_CESD7_1992 = n(), # calculates the sample size for the CESD7_1992 variable
            mean_CESD7_1994 = mean(CESD7_1994, na.rm = TRUE), # calculates the mean for the CESD7_1994 variable
            sd_CESD7_1994 = sd(CESD7_1994, na.rm = TRUE),  # calculates the standard deviation for the CESD7_1994 variable
            n_CESD7_1994 = n(), # calculates the sample size for the CESD7_1994 variable
            mean_CESD20_1992 = mean(CESD20_1992, na.rm = TRUE), # calculates the mean for the CESD20_1992 variable
            sd_CESD20_1992 = sd(CESD20_1992, na.rm = TRUE),  # calculates the standard deviation for the CESD20_1992 variable
            n_CESD20_1992 = n()) %>%   # calculates the sample size for the CESD20_1992 variable
  ungroup() # remove grouping from the data frame

# Display the resulting data frame as a nice table using the gt package
gt(df_data_RACE, title = "Depression and Delinquency means divided by Race")


summary(df_data_gen1)

```


```{r data visualization} 
#Data visualization
#sex, minoirity/race, age.
#correlation table


df_data_gen1 %>%
  select(DELIN_AVERAGE, SEX, AGE_1979, CESD7_1992, CESD20_1992, CESD7_1994, MINORITY) %>%
tbl_summary(
    label= list(
  DELIN_AVERAGE ~ "Delinquency Average",
  AGE_1979 ~ "Age" ))


ggplot(df_data_gen1, aes(x =df_data_gen1$MINORITY)) + 
  geom_bar(fill = "cornflowerblue", 
           color="black") +
  labs(x = "Race", 
       y = "Frequency", 
       title = "Participants by Race") +
   theme_minimal()

ggplot(df_data_gen1, aes(x =AGE_1979)) + 
  geom_bar(fill = "Light Blue", 
           color="black") +
  labs(x = "Age", 
       y = "Participant Amount", 
       title = "Participant Age") +
  scale_x_continuous(breaks = seq(14, 23)) +
   theme_minimal()

#have female be 0 and male be 1
   ggplot(df_data_gen1, aes(x = SEX))+
       geom_bar(fill = "cornflowerblue", 
           color="black") +
  labs(x = "Gender", 
       y = "Frequency", 
       title = "Participants by Gender",
       subtitle = "0=female, 1=male")

   
#To convert the given bar plot code into a violin plot that shows the distribution of DELIN_AVERAGE grouped #by MINORITY, you can make the following changes to the code:

#1. Replace `geom_bar` with `geom_violin`.
#2. Change the aesthetics `aes` to include `y = DELIN_AVERAGE`.
#3. Modify the `labs` to update the y-axis label.
#4. Update the fill and color options to your preference.

#Here's the modified code:


ggplot(df_data_gen1, aes(x = MINORITY, y = DELIN_AVERAGE)) + 
  geom_violin(fill = "cornflowerblue", 
              color = "black") +
  labs(x = "Race", 
       y = "DELIN_AVERAGE", 
       title = "Distribution of DELIN_AVERAGE by Race") +
  theme_minimal()



ggplot(df_data_gen1, aes(x = SEX, y = DELIN_AVERAGE)) + 
  geom_violin(fill = "cornflowerblue", 
              color = "black") +
  labs(x = "SEX", 
       y = "DELIN_AVERAGE", 
       title = "Distribution of DELIN_AVERAGE by Sex") +
  theme_minimal()

ggplot(df_data_gen1, aes(x = AGE_1979, y = DELIN_AVERAGE)) + 
  geom_violin(fill = "cornflowerblue", 
              color = "black") +
  labs(x = "AGE_1979", 
       y = "DELIN_AVERAGE", 
       title = "Distribution of DELIN_AVERAGE by AGE_1979") +
  theme_minimal()
```



## Summarize

# Data Linking

```{r kinship, warning=FALSE,include=FALSE}

#link dyads

dsLinks <- Links79PairExpanded
dsLinks$R <- dsLinks$RFull
dsLinking <- dsLinks

outcomeNames <- c("MINORITY",
                  "SEX",
                  "CESD20_1992",
                  "DELIN_AVERAGE",
                  "CESD7_1994",
                  "CESD7_1992",
                  "AGE_1979")

# only full full sibs
dsSingle_G1 <- CreatePairLinksSingleEntered(
  outcomeDataset = df_data_gen1,
  linksPairDataset = dsLinking[dsLinking$RelationshipPath == "Gen1Housemates" &
                                 dsLinking$RFull == .5,],
  outcomeNames = outcomeNames,
  subject1Qualifier = "_S1",
  subject2Qualifier = "_S2"
)




set.seed(18)



df_dyad_delinq <- dsSingle_G1 %>%
  drop_na(contains("DELIN_AVERAGE")) %>%
  discord_data(
    outcome = "DELIN_AVERAGE",
    predictors = c("CESD7_1992",
                   "CESD20_1992",
                   "CESD7_1994",
                   "AGE_1979"),
    id = "ExtendedID",
    sex = "SEX",
    race = "MINORITY",
    pair_identifiers = c("_S1", "_S2"),
    demographics = "both"
  ) %>%
  mutate(
    # sex is 0 for girl; and 1 for boy
    SEX_all4 = case_when(
      # boyboy
      SEX_1 == 1 &
        SEX_2 == 1 ~ "BoyBoy",
      # girlboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "GirlBoy",
      #opposite
      SEX_1 == 1 &
        SEX_2 == 0 ~ "BoyGirl",
      #girgirl
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    ),
    #opposite
    SEX_all3 = case_when(
      SEX_1 == 1 & SEX_2 == 1 ~ "BoyBoy",
      # boyboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "mixed",
      # girlboy
      SEX_1 == 1 &
        SEX_2 == 0 ~ "mixed",
      #opposite
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    )
  )



df_dyad_CESD20_1992 <- dsSingle_G1 %>%
  drop_na(contains("CESD20_1992")) %>%
  discord_data(
    outcome = "CESD20_1992",
    predictors = c("CESD7_1992",
                   "DELIN_AVERAGE",
                   "CESD7_1994",
                   "AGE_1979"),
    id = "ExtendedID",
    sex = "SEX",
    race = "MINORITY",
    pair_identifiers = c("_S1", "_S2"),
    demographics = "both"
  ) %>%
  mutate(
    # sex is 0 for girl; and 1 for boy
    SEX_all4 = case_when(
      # boyboy
      SEX_1 == 1 &
        SEX_2 == 1 ~ "BoyBoy",
      # girlboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "GirlBoy",
      #opposite
      SEX_1 == 1 &
        SEX_2 == 0 ~ "BoyGirl",
      #girgirl
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    ),
    #opposite
    SEX_all3 = case_when(
      SEX_1 == 1 & SEX_2 == 1 ~ "BoyBoy",
      # boyboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "mixed",
      # girlboy
      SEX_1 == 1 &
        SEX_2 == 0 ~ "mixed",
      #opposite
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    )
  )

df_dyad_CESD7_1992 <- dsSingle_G1 %>%
  drop_na(contains("CESD7_1992")) %>%
  discord_data(
    outcome = "CESD7_1992",
    predictors = c("CESD20_1992",
                   "DELIN_AVERAGE",
                   "CESD7_1994",
                   "AGE_1979"),
    id = "ExtendedID",
    sex = "SEX",
    race = "MINORITY",
    pair_identifiers = c("_S1", "_S2"),
    demographics = "both"
  ) %>%
  mutate(
    # sex is 0 for girl; and 1 for boy
    SEX_all4 = case_when(
      # boyboy
      SEX_1 == 1 &
        SEX_2 == 1 ~ "BoyBoy",
      # girlboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "GirlBoy",
      #opposite
      SEX_1 == 1 &
        SEX_2 == 0 ~ "BoyGirl",
      #girgirl
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    ),
    #opposite
    SEX_all3 = case_when(
      SEX_1 == 1 & SEX_2 == 1 ~ "BoyBoy",
      # boyboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "mixed",
      # girlboy
      SEX_1 == 1 &
        SEX_2 == 0 ~ "mixed",
      #opposite
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    )
  )


df_dyad_CESD7_1994 <- dsSingle_G1 %>%
  drop_na(contains("CESD7_1994")) %>%
  discord_data(
    outcome = "CESD7_1994",
    predictors = c("CESD20_1992",
                   "DELIN_AVERAGE",
                   "CESD7_1992",
                   "AGE_1979"),
    id = "ExtendedID",
    sex = "SEX",
    race = "MINORITY",
    pair_identifiers = c("_S1", "_S2"),
    demographics = "both"
  ) %>%
  mutate(
    # sex is 0 for girl; and 1 for boy
    SEX_all4 = case_when(
      # boyboy
      SEX_1 == 1 &
        SEX_2 == 1 ~ "BoyBoy",
      # girlboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "GirlBoy",
      #opposite
      SEX_1 == 1 &
        SEX_2 == 0 ~ "BoyGirl",
      #girgirl
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    ),
    #opposite
    SEX_all3 = case_when(
      SEX_1 == 1 & SEX_2 == 1 ~ "BoyBoy",
      # boyboy
      SEX_1 == 0 &
        SEX_2 == 1 ~ "mixed",
      # girlboy
      SEX_1 == 1 &
        SEX_2 == 0 ~ "mixed",
      #opposite
      SEX_1 == 0 &
        SEX_2 == 0 ~ "GirlGirl"
    )
  )

```

# classic regression

```{r classic}
#naming notation

# lm is linear model
# og is classic regression
# xdelin means that delinquency is the predictor
# ydelin means that delinguency is the outcome

df_classic <- df_data_gen1

lm_og_ydelin_xCESD7_1992 <- lm(DELIN_AVERAGE ~
                     CESD7_1992 +
                     as.factor(SEX) +
                     AGE_1979+
                     as.factor(MINORITY) ,
                   data = df_classic)



lm_og_ydelin_xCESD20_1992 <- lm(DELIN_AVERAGE ~
                      CESD20_1992 +
                      as.factor(SEX) +
                        AGE_1979+
                      as.factor(MINORITY),
                    data = df_classic)



lm_og_ydelin_xCESD7_1994 <- lm(DELIN_AVERAGE ~
                     CESD7_1994 +
                     as.factor(SEX) +
                       AGE_1979+
                     as.factor(MINORITY) ,
                   data = df_classic)



lm_og_xdelin_yCESD7_1992 <- lm(CESD7_1992 ~
                     DELIN_AVERAGE +
                     as.factor(SEX) +
                       AGE_1979+
                     as.factor(MINORITY) ,
                   data = df_classic)

lm_og_xdelin_yCESD20_1992 <- lm(CESD20_1992 ~
                     DELIN_AVERAGE +
                      as.factor(SEX) +
                       AGE_1979+
                      as.factor(MINORITY),
                    data = df_classic)


lm_og_xdelin_yCESD7_1994 <- lm(CESD7_1994 ~
                      DELIN_AVERAGE+
                     as.factor(SEX) +
                       AGE_1979+
                     as.factor(MINORITY) ,
                   data = df_classic)




```

# between-family

```{r betweenfamily}

#naming notation

# lm is linear model
# btw is between family
# xdelin means that delinquency is the predictor
# ydelin means that delinguency is the outcome
# sex4 means that there are 4 sex categories
# sex3 means that there are 3

df_btw <- df_dyad_delinq



#note: you only need s1 race because they're determined by the mother
lm_btw_ydelin_sex4_xCESD7_1992 <- lm(DELIN_AVERAGE_mean ~ 
                             CESD7_1992_mean +
                             SEX_all4 +
                              AGE_1979_mean+
                           #  factor(SEX_1)+factor(SEX_2)+
                            as.factor(MINORITY_1),
                           data = df_btw)

lm_btw_ydelin_sex3_xCESD7_1992 <- lm(DELIN_AVERAGE_mean ~ 
                             CESD7_1992_mean +
                            SEX_all3 +
                           #  factor(SEX_1)+factor(SEX_2)+
                             AGE_1979_mean+
                            as.factor(MINORITY_1),
                           data = df_btw)


lm_btw_ydelin_sex3_xCESD20_1992 <- lm(DELIN_AVERAGE_mean ~ 
                             CESD20_1992_mean +
                            SEX_all3 +
                              AGE_1979_mean+
                           #  factor(SEX_1)+factor(SEX_2)+
                            as.factor(MINORITY_1), 
                           data = df_btw)

lm_btw_xdelin_sex3_yCESD20_1992 <- lm(CESD20_1992_mean ~ 
                             DELIN_AVERAGE_mean +
                            SEX_all3 +
                              AGE_1979_mean+
                           #  factor(SEX_1)+factor(SEX_2)+
                            as.factor(MINORITY_1), 
                           data = df_btw)

summary(lm_btw_ydelin_sex3_xCESD20_1992)
summary(lm_btw_xdelin_sex3_yCESD20_1992)

lm_btw_ydelin_sex4_xCESD20_1992 <- lm(DELIN_AVERAGE_mean ~ 
                             CESD20_1992_mean +
                            SEX_all4 +
                              AGE_1979_mean+
                           #  factor(SEX_1)+factor(SEX_2)+
                            as.factor(MINORITY_1), 
                           data = df_btw)

lm_btw_ydelin_sex4_xCESD7_1994 <- lm(DELIN_AVERAGE_mean ~ 
                            CESD7_1994_mean +
                            SEX_all4 +
                              AGE_1979_mean+
                           #  factor(SEX_1)+factor(SEX_2)+
                            as.factor(MINORITY_1),
                           data = df_btw)

lm_btw_ydelin_sex3_xCESD7_1994 <- lm(DELIN_AVERAGE_mean ~ 
                            CESD7_1994_mean +
                            SEX_all3 +
                            AGE_1979_mean+
                           #  factor(SEX_1)+factor(SEX_2)+
                            as.factor(MINORITY_1),
                           data = df_btw)

```

# within-family analyses

```{r discordmodels}


lm_discord_ydelin_sex4_xCESD7_1992 <- lm(data=df_dyad_delinq,
                                DELIN_AVERAGE_diff ~ 
                                DELIN_AVERAGE_mean+
                                CESD7_1992_diff + 
                                CESD7_1992_mean+
                                  AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all4 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_ydelin_sex3_xCESD7_1992 <- lm(data=df_dyad_delinq,
                                DELIN_AVERAGE_diff ~ 
                                DELIN_AVERAGE_mean+
                                CESD7_1992_diff + 
                                CESD7_1992_mean+
                                     AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all3 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_ydelin_sex3_xCESD20_1992 <- lm(data=df_dyad_delinq,
                              DELIN_AVERAGE_diff ~ 
                                DELIN_AVERAGE_mean +
                                CESD20_1992_diff + 
                                CESD20_1992_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all3 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_ydelin_sex4_xCESD20_1992 <- lm(data=df_dyad_delinq,
                              DELIN_AVERAGE_diff ~ 
                                DELIN_AVERAGE_mean +
                                CESD20_1992_diff + 
                                CESD20_1992_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all4 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )


lm_discord_ydelin_sex3_xCESD7_1994 <- lm(data=df_dyad_delinq,
                              DELIN_AVERAGE_diff ~ 
                                DELIN_AVERAGE_mean +
                                CESD7_1994_diff + 
                                CESD7_1994_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all3 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_ydelin_sex4_xCESD7_1994 <- lm(data=df_dyad_delinq,
                              DELIN_AVERAGE_diff ~ 
                                DELIN_AVERAGE_mean +
                                CESD7_1994_diff + 
                                CESD7_1994_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all4 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )



df_dyad<-df_dyad_CESD7_1992

lm_discord_xdelin_sex4_yCESD7_1992 <- lm(data=df_dyad,
                                CESD7_1992_diff ~ 
                                DELIN_AVERAGE_mean+
                                DELIN_AVERAGE_diff + 
                                CESD7_1992_mean+
                                     AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all4 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_xdelin_sex3_yCESD7_1992 <- lm(data=df_dyad,
                                CESD7_1992_diff ~ 
                                DELIN_AVERAGE_mean+
                                DELIN_AVERAGE_diff + 
                                CESD7_1992_mean+
                                     AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all3 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )
df_dyad<-df_dyad_CESD20_1992

lm_discord_xdelin_sex3_yCESD20_1992 <- lm(data=df_dyad,
                              CESD20_1992_diff ~ 
                                DELIN_AVERAGE_mean +
                                DELIN_AVERAGE_diff + 
                                CESD20_1992_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all3 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_xdelin_sex4_yCESD20_1992 <- lm(data=df_dyad,
                              CESD20_1992_diff ~ 
                                DELIN_AVERAGE_mean +
                                DELIN_AVERAGE_diff + 
                                CESD20_1992_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all4 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )
df_dyad<-df_dyad_CESD7_1994

lm_discord_xdelin_sex3_yCESD7_1994 <- lm(data=df_dyad,
                              CESD7_1994_diff ~ 
                                DELIN_AVERAGE_mean +
                                DELIN_AVERAGE_diff + 
                                CESD7_1994_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all3 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )

lm_discord_xdelin_sex4_yCESD7_1994 <- lm(data=df_dyad,
                              CESD7_1994_diff ~ 
                                DELIN_AVERAGE_mean +
                                DELIN_AVERAGE_diff + 
                                CESD7_1994_mean +
                                   AGE_1979_diff+
                                  AGE_1979_mean+
                                SEX_all4 +
                              #  factor(SEX_1)+factor(SEX_2)+
                                factor(MINORITY_1) )
```

## Results

### raw

```{r raw}

#classical regression

lm_og_ydelin_xCESD7_1992  %>%  
prettify_regression_results(standardize=TRUE)
lm_og_ydelin_xCESD20_1992  %>%  
prettify_regression_results(standardize=TRUE)

lm_og_ydelin_xCESD7_1994  %>%  
prettify_regression_results(standardize=TRUE)
```

### Btween

```{r raw btw}

#between family 

lm_btw_ydelin_sex3_xCESD7_1992  %>%  
prettify_regression_results(standardize=TRUE)

#between family

lm_btw_ydelin_sex3_xCESD20_1992  %>%  
prettify_regression_results(standardize=TRUE)

lm_btw_ydelin_sex3_xCESD7_1994  %>%  
prettify_regression_results(standardize=TRUE)

lm_btw_xdelin_sex3_yCESD20_1992 %>%  
prettify_regression_results(standardize=TRUE)

```


### Discordant

#### CESD_7\_1992

```{r discord}

lm_discord_ydelin_sex3_xCESD7_1992  %>%  
prettify_regression_results(standardize=TRUE)


lm_discord_xdelin_sex3_yCESD7_1992  %>%  
prettify_regression_results(standardize=TRUE)

```
CESD7_1994

```{r raw3}

lm_discord_ydelin_sex3_xCESD7_1994 %>%  
prettify_regression_results(standardize=TRUE)

lm_discord_xdelin_sex3_yCESD7_1994 %>%  
prettify_regression_results(standardize=TRUE)

```

#CESD_20_1992

The CESD_20 depression difference was not significant.
The CESD_20 depression score did not predict criminal behavior.
This means that the difference was not meaningful above or beyond the other sources of variance.
Our hypothesis was not supported.
Increased depression did not lead to more crime in adolescents.
The variance in criminal behavior explained by the regression model is 61.48% f(6, 2950) = 787.4, p = \< 2.2e-16.
The Sex Sibling 1 was significant t(2950) = 5.141, p= 2.91e-07.
Sex Sibling 2 was signifcant t(2950) = -6.513, p = 8.62e-11.



```{r raw5}

lm_discord_ydelin_sex3_xCESD20_1992 %>% 
prettify_regression_results(standardize=TRUE)


lm_discord_xdelin_sex3_yCESD20_1992 %>%  
prettify_regression_results(standardize=TRUE)
```

# Results

We examined the relationship between depression and crime in three distinct stages.
First, we looked at the overall relationship without accounting for family structure. Second, we looked at the between-family relationships. between-family analyses report the relationships between the within-family average AFI and various measures of ability. Lastly we looked at the within-family relationships. Throughout, this process we also examined whether gender, race, and age impacted these dynamics.

## Unadjusted

Unadjusted univariate and bivariate descriptive statistics were consist with what others have found in this literature. See tables X and Y. Of note, 

```{r og dep predict delinq}
library(lm.beta)
lm_result <- lm.beta(lm_og_ydelin_xCESD20_1992)
print(lm_result)
summary(lm_result)
result_coef <- coef(lm_result)
writeup_btw<-summary(lm_result)
result_r2<-round(writeup_btw$adj.r.squared,4)
result_f<-round(writeup_btw$fstatistic['value'],4)
result_ndf<-round(writeup_btw$fstatistic['numdf'],0)
result_ddf<-round(writeup_btw$fstatistic['dendf'],0)
result_n <-length(writeup_btw$residuals)
result_predictor_mean <- round(result_coef['CESD20_1992'],4)
result_age_mean <- round(result_coef['AGE_1979'],4)
result_bg <- round(result_coef['as.factor(SEX)1'],4)
result_minority <- round(result_coef['as.factor(MINORITY)nonblackhispanic'],4)
```
Averages of CESD-20 were used to predict averages of delinquency. Table X presents the unstandardized regression results (n =`r result_n`). The F-statistic (`r result_f`,df1=`r result_ndf`,df2=`r result_ddf`) was significant, and the adjusted R$^{2}$ was `r result_r2`. A standard-deviation increase in average depression predicted a statistically significant increase of `r result_predictor_mean` standard deviations in the average delinquency score.  As expected, gender composition, age, and minority status were also predictive. Being female was predictive of having more average delinquency (`r result_bg`). Similarly, older age was predictive of lower levels of delinquency (`r result_age_mean`). Being in the nonminority group was predictive of `r result_minority` standard deviations for average of the delinquency measure.


## Between-Family

We examined the between-family results. We tested whether the sibling average for delinquency was associated with sibling averages in depression. We examined this relationship in both directions, in other words we examined whether delinquency predicted depression or whether depression predicted delinquency. Results, as expected, were similar. Accordingly we report results looking at <direction of interest>, and direct interested readers to the appendix. Similarly, we focused on the CES-D 20 as our measure of depression, but include results for the other two variants in the appendix. We discuss the variants of these models, when the results differ.

### Depression predicting deliquency
```{r dep predict delinq}

library(lm.beta)
lm_result <- lm.beta(lm_btw_ydelin_sex3_xCESD20_1992)
print(lm_result)
summary(lm_result)
result_coef <- coef(lm_result)
writeup_btw<-summary(lm_result)
result_r2<-round(writeup_btw$adj.r.squared,4)
result_f<-round(writeup_btw$fstatistic['value'],4)
result_ndf<-round(writeup_btw$fstatistic['numdf'],0)
result_ddf<-round(writeup_btw$fstatistic['dendf'],0)
result_n <-length(writeup_btw$residuals)
result_predictor_mean <- round(result_coef['CESD20_1992_mean'],4)
result_age_mean <- round(result_coef['AGE_1979_mean'],4)
result_gg <- round(result_coef['SEX_all3GirlGirl'],4)
result_bg <- round(result_coef['SEX_all3mixed'],4)
result_minority <- round(result_coef['as.factor(MINORITY_1)nonblackhispanic'],4)
```


Sibling averages of CESD-20 were used to predict sibling averages of delinquency. Table X presents the unstandardized regression results (n =`r result_n`). The F-statistic (`r result_f`,df1=`r result_ndf`,df2=`r result_ddf`) was significant, and the adjusted R$^{2}$ was `r result_r2`. A standard-deviation increase in the siblings average depression predicted a statistically significant increase of `r result_predictor_mean` standard deviations in the average of the siblings' average delinquency score.  As expected, gender composition, age, and minority status were also predictive. Relative to the reference group of brothers, families with one (`r result_bg`) or two (`r result_gg`) sisters had lower averages in delinquency. Similarly, older sibling pairs were predictive of lower levels of delinquency (`r result_age_mean`). Relative to the reference group (of respondents classified as either Black or Hispanic), had `r result_minority` standard deviations for average of the siblings' delinquency measure.

### Deliquency predicting depression
```{r delin pred dep2}

library(lm.beta)

lm_result <- lm.beta(lm_btw_xdelin_sex3_yCESD20_1992)
print(lm_result)
summary(lm_result)
result_coef <- coef(lm_result)
writeup_btw<-summary(lm_result)
result_r2<-round(writeup_btw$adj.r.squared,4)
result_f<-round(writeup_btw$fstatistic['value'],4)
result_ndf<-round(writeup_btw$fstatistic['numdf'],0)
result_ddf<-round(writeup_btw$fstatistic['dendf'],0)
result_n <-length(writeup_btw$residuals)
result_predictor_mean <- round(result_coef['DELIN_AVERAGE_mean'],4)
result_age_mean <- round(result_coef['AGE_1979_mean'],4)
result_gg <- round(result_coef['SEX_all3GirlGirl'],4)
result_bg <- round(result_coef['SEX_all3mixed'],4)
result_minority <- round(result_coef['as.factor(MINORITY_1)nonblackhispanic'],4)
```

Sibling averages of delinquency were used to predict sibling averages of depression. Table X presents the unstandardized regression results (n =`r result_n`). The F-statistic (`r result_f`,df1=`r result_ndf`,df2=`r result_ddf`) was significant, and the adjusted R$^{2}$ was `r result_r2`. A standard-deviation increase in the siblings average delinquency predicted a statistically significant increase of `r result_predictor_mean` standard deviations in the average of the siblings' depression. As expected, gender composition and minority status were also predictive, whereas age was not. Relative to the reference group of brothers, families with one (`r result_bg`) or two (`r result_gg`) sisters had higher averages in depression. Relative to the reference group, non-Black non-Hispanic families had higher family levels of depression (b = `r result_minority`).


## Sibling Difference Score

We then implemented within-family versions of the between-family results we presented in the previous subsection. Specifically, we modeled the relationship of sibling differences in depression and differences in delinquency. We modeled them in both directions, while controlling for the means of outcomes and predictors. Just as we did in the between-family analyses, we applied these models to the other measures of depression. They are available in the appendix, but we make note of any meaningful differences.

### Deliquency predicting depression
```{r deli predict depr}

library(lm.beta)
lm_result <- lm.beta(lm_discord_xdelin_sex3_yCESD20_1992)
print(lm_result)
summary(lm_result)
result_coef <- coef(lm_result)
writeup_wth <- summary(lm_result)
result_r2 <- round(writeup_wth$adj.r.squared,4)
result_f <- round(writeup_wth$fstatistic['value'],4)
result_ndf <- round(writeup_wth$fstatistic['numdf'],0)
result_ddf <- round(writeup_wth$fstatistic['dendf'],0)
result_n <- length(writeup_wth$residuals)
result_outcome_mean <- round(result_coef['CESD20_1992_diff'],4)
result_predictor_mean <- round(result_coef['DELIN_AVERAGE_mean'],4)
result_predictor_diff <- round(result_coef['DELIN_AVERAGE_diff'],4)
result_age_mean <- round(result_coef['AGE_1979_mean'],4)
result_age_diff <- round(result_coef['AGE_1979_diff'],4)
result_gg <- round(result_coef['SEX_all3GirlGirl'],4)
result_bg <- round(result_coef['SEX_all3mixed'],4)
result_minority <- round(result_coef['as.factor(MINORITY_1)nonblackhispanic'],4)
```

### Deliquency predicting depression
### 

Sibling differences in delinquency were used to predict differences in depression, controlling for family level gender composition, family minority status, and sibling averages of both delinquency and depression. Table X presents the unstandardized regression results (n =`r result_n`). The F-statistic (`r result_f`,df1=`r result_ndf`,df2=`r result_ddf`) was significant, and the adjusted R$^{2}$ was `r result_r2`. A standard-deviation increase in the average delinquency predicted a statistically significant increase of `r result_predictor_mean` standard deviations in the average of the siblings' depression. Relative to the reference group, non-Black non-Hispanic families had higher family levels of depression (b = `r result_minority`). Similarly, older sibling pairs were predictive of lower levels of depression (`r result_age_diff`).

##depression predicting delinquency

```{r dep predict deli}

library(lm.beta)
lm_result <- lm.beta(lm_discord_ydelin_sex3_xCESD20_1992)
print(lm_result)
summary(lm_result)
result_coef <- coef(lm_result)
writeup_wth <- summary(lm_result)
result_r2 <- round(writeup_wth$adj.r.squared,4)
result_f <- round(writeup_wth$fstatistic['value'],4)
result_ndf <- round(writeup_wth$fstatistic['numdf'],0)
result_ddf <- round(writeup_wth$fstatistic['dendf'],0)
result_n <- length(writeup_wth$residuals)
result_outcome_mean <- round(result_coef['DELIN_AVERAGE_diff'],4)
result_predictor_mean <- round(result_coef['CESD20_1992_mean'],4)
result_predictor_diff <- round(result_coef['CESD20_1992_diff'],4)
result_age_mean <- round(result_coef['AGE_1979_mean'],4)
result_age_diff <- round(result_coef['AGE_1979_diff'],4)
result_gg <- round(result_coef['SEX_all3GirlGirl'],4)
result_bg <- round(result_coef['SEX_all3mixed'],4)
result_minority <- round(result_coef['as.factor(MINORITY_1)nonblackhispanic'],4)
```

## depression predicting delinquency

Sibling differences in depression were used to predict differences in delinquency, controlling for family level gender composition, family minority status, and sibling averages of both delinquency and depression. Table X presents the unstandardized regression results (n =`r result_n`). The F-statistic (`r result_f`,df1=`r result_ndf`,df2=`r result_ddf`) was significant, and the adjusted R$^{2}$ was `r result_r2`. A standard-deviation increase in the average delinquency predicted a statistically significant increase of `r result_predictor_mean` standard deviations in the average of the siblings' depression. Similarly, older sibling pairs were predictive of lower levels of delinquency (`r result_age_diff`). Relative to the reference group of brothers, families with one (`r result_bg`) sister had higher averages in delinquency.

All three models reveal similar results. Averages of delinquency were significant predictors of differences in delinquency (p $< .01$). A one standard deviation increase in the sibling average of delinquency predicted a `r result_outcome_mean` increase in average delinquency difference, controlling for all over variables in the model. (We don't encourage people to interpret these coefficients deeply).


